### 2017-05-08 tex/fsm_two.tex
FSM graphs and process pseudocode
This closes #29 and #28.

The graph representation for the e finite state machines
as described in `players.py`.

Also for representing the process two pieces of pseudocode.
One for the data collection and one for the moran process.

The pseudocodes are in a tex file, not sure where you would like
them.

Update gitignore to ignore pdf of tex files.

Ashlock's fortress strategies for comparison are also included.n

\documentclass{standalone}

\usepackage{tikz}
\usepackage{standalone}
\usetikzlibrary{calc}

\begin{document}
    \begin{tikzpicture}

    \tikzstyle{state}=[minimum width=0.8cm, font=\boldmath];
    \node[circle, draw=black, thick] (1) at (0, 0) [state] {$2$};
	\node[circle, draw=black, thick] (4) at (4, 0) [state] {$5$};

	\node[circle, draw=black, thick] (5) at (-2, -1) [state] {$6$};
    \node[circle, draw=black, thick] (6) at (2, -1) [state] {$7$};
    \node[circle, draw=black, thick] (3) at ($(5)+(-1,-2)$) [state] {$4$};
    \node[circle, draw=black, thick] (12) at ($(5)+(0,-4)$) [state] {$13$};
    \node[circle, draw=black, thick] (8) at ($(12)+(0,-2)$) [state] {$9$};
    \node[circle, draw=black, thick] (9) at ($(8)+(-2,0)$) [state] {$10$};
    \node[circle, draw=black, thick] (0) at ($(8)+(0,-2)$) [state] {$1$};

    \node[circle, draw=black, thick] (2) at ($(6)+(0,-2)$) [state] {$3$};
    \node[circle, draw=black, thick] (14) at ($(2)+(2,0)$) [state] {$15$};
    \node[circle, draw=black, thick] (7) at ($(2)+(0,-2)$) [state] {$8$};
    \node[circle, draw=black, thick] (10) at ($(14)+(0,-2)$) [state] {$11$};
    \node[circle, draw=black, thick] (13) at ($(7)+(0,-2)$) [state] {$14$};
    \node[circle, draw=black, thick] (11) at ($(10)+(0,-2)$) [state] {$12$};
    \node[circle, draw=black, thick] (15) at ($(11)+(0,-2)$) [state] {$16$};

    \coordinate[below of=0] (s);

    \draw (s) edge[out=90, in=-90, ->, thick] node [left] {$C$} (0);

    \draw (0) edge[out=0, in=-90, ->, thick] node [below right] {$C/D$} (13);
    \draw (0) edge[out=45, in=0, ->, thick] node [above, rotate=90] {$D/D$} (12);

    \draw (1) edge[out=-90, in=0, ->, thick] node [above, yshift=1.3cm, xshift=0.3cm] {$C/D$} (3);
    \draw (1) edge[out=0, in=180, ->, thick] node [above] {$D/D$} (4);

    \draw (2) edge[out=-45, in=-135, ->, thick] node [above] {$C/D$} (14);
    \draw (2) edge[out=180, in=90, ->, thick] node [above, yshift=0.8cm, xshift=3cm] {$D/D$} (9);

    \draw (3) edge[out=180, in=-135, ->, thick] node [left, yshift=2cm, xshift=-0.1cm] {$C/C$} (0);
    \draw (3) edge[out=135, in=180, ->, thick] node [above, yshift=0.2cm] {$D/C$} (1);

    \draw (4) edge[out=90, in=90, ->, thick] node [above] {$C/D$} (1);
    \draw (4) edge[out=-90, in=0, ->, thick] node [right] {$D/D$} (2);

    \draw (5) edge[out=-90, in=90, ->, thick] node [right, yshift=1cm] {$C/C$} (12);
    \draw (5) edge[out=0, in=180, ->, thick] node [below right] {$D/C$} (6);

    \draw (6) edge[out=135, in=-45, ->, thick] node [right, xshift=1cm] {$C/C$} (1);
    \draw (6) edge[out=-45, in=135, ->, thick] node [above left, yshift=-0.5cm] {$D/D$} (14);

    \draw (7) edge[out=135, in=45, ->, thick] node [above] {$C/D$} (12);
    \draw (7) edge[out=90, in=-90, ->, thick] node [left] {$D/D$} (2);

    \draw (8) edge[out=90, in=-135, ->, thick] node [above right] {$C/D$} (7);
    \draw (8) edge[out=180, in=0, ->, thick] node [above] {$D/D$} (9);

    \draw (9) edge[out=-45, in=-135, ->, thick] node [below] {$C/D$} (8);
    \draw (9) edge[out=-90, in=180, ->, thick] node [right] {$D/D$} (0);

    \draw (10) edge[out=180, in=-60, ->, thick] node [below, yshift=-0.5cm, xshift=0.4cm] {$C/C$} (2);
    \draw (10) edge[out=0, in=0, ->, thick] node [right] {$D/C$} (15);

    \draw (11) edge[out=90, in=-45, ->, thick] node [below right, xshift=0.7cm] {$C/D$} (7);
    \draw (11) edge[out=180, in=0, ->, thick] node [below] {$D/D$} (13);

    \draw (12) edge[out=135, in=-45, ->, thick] node [above left] {$C/C$} (3);
    \draw (12) edge[out=-90, in=135, ->, thick] node [left] {$D/D$} (8);

    \draw (13) edge[out=90, in=-90, ->, thick] node [left] {$C/C$} (7);
    \draw (13) edge[out=45, in=-135, ->, thick] node [below, yshift=-0.2cm, rotate=45] {$D/D$} (10);

    \draw (14) edge[out=-90, in=90, ->, thick] node [right] {$C/D$} (10);
    \draw (14) edge[out=-100, in=45, ->, thick] node [above] {$D/D$} (7);

    \draw (15) edge[out=-135, in=180, ->, thick, loop] node [left] {$C/C$} (15);
    \draw (15) edge[out=90, in=-90, ->, thick] node [left] {$D/D$} (11);

    \end{tikzpicture}
\end{document}