### 2017-05-10 tex/list_of_players.tex
Adjust nb and rerun images.

\item $\phi$
\item $\pi$
\item $e$
\item ALLCorALLD
\item Adaptive
\item Adaptive Pavlov 2006
\item Adaptive Pavlov 2011
\item Adaptive Tit For Tat: 0.5
\item Aggravater
\item Alternator
\item Alternator Hunter
\item Anti Tit For Tat
\item AntiCycler
\item Appeaser
\item Arrogant QLearner
\item Average Copier
\item Better and Better
\item Bully
\item Calculator
\item Cautious QLearner
\item CollectiveStrategy
(\textbf{CS})\item Contrite Tit For Tat
(\textbf{CTfT})\item Cooperator
\item Cooperator Hunter
\item Cycle Hunter
\item Cycler CCCCCD
\item Cycler CCCD
\item Cycler CCCDCD
\item Cycler CCD
\item Cycler DC
\item Cycler DDC
\item Davis: 10
\item Defector
\item Defector Hunter
\item Desperate
\item Doubler
\item EasyGo
\item Eatherley
\item Eventual Cycle Hunter
\item Evolved ANN
\item Evolved ANN 5
\item Evolved ANN 5 Noise 05
\item Evolved FSM 16
\item Evolved FSM 16 Noise 05
\item Evolved FSM 4
\item Evolved HMM 5
\item EvolvedLookerUp1\_1\_1
\item EvolvedLookerUp2\_2\_2
\item FSM Player: [(0, 'C', 0, 'C'), (0, 'D', 3, 'C'), (1, 'C', 5, 'D'), (1, 'D', 0, 'C'), (2, 'C', 3, 'C'), (2, 'D', 2, 'D'), (3, 'C', 4, 'D'), (3, 'D', 6, 'D'), (4, 'C', 3, 'C'), (4, 'D', 1, 'D'), (5, 'C', 6, 'C'), (5, 'D', 3, 'D'), (6, 'C', 6, 'D'), (6, 'D', 6, 'D'), (7, 'C', 7, 'D'), (7, 'D', 5, 'C')], 0, C
(\textbf{Trained FSM 1})\item FSM Player: [(0, 'C', 13, 'D'), (0, 'D', 12, 'D'), (1, 'C', 3, 'D'), (1, 'D', 4, 'D'), (2, 'C', 14, 'D'), (2, 'D', 9, 'D'), (3, 'C', 0, 'C'), (3, 'D', 1, 'D'), (4, 'C', 1, 'D'), (4, 'D', 2, 'D'), (5, 'C', 12, 'C'), (5, 'D', 6, 'C'), (6, 'C', 1, 'C'), (6, 'D', 14, 'D'), (7, 'C', 12, 'D'), (7, 'D', 2, 'D'), (8, 'C', 7, 'D'), (8, 'D', 9, 'D'), (9, 'C', 8, 'D'), (9, 'D', 0, 'D'), (10, 'C', 2, 'C'), (10, 'D', 15, 'C'), (11, 'C', 7, 'D'), (11, 'D', 13, 'D'), (12, 'C', 3, 'C'), (12, 'D', 8, 'D'), (13, 'C', 7, 'C'), (13, 'D', 10, 'D'), (14, 'C', 10, 'D'), (14, 'D', 7, 'D'), (15, 'C', 15, 'C'), (15, 'D', 11, 'D')], 0, C
(\textbf{Trained FSM 2})\item FSM Player: [(0, 'C', 7, 'C'), (0, 'D', 1, 'C'), (1, 'C', 11, 'D'), (1, 'D', 11, 'D'), (2, 'C', 8, 'D'), (2, 'D', 8, 'C'), (3, 'C', 3, 'C'), (3, 'D', 12, 'D'), (4, 'C', 6, 'C'), (4, 'D', 3, 'C'), (5, 'C', 11, 'C'), (5, 'D', 8, 'D'), (6, 'C', 13, 'D'), (6, 'D', 14, 'C'), (7, 'C', 4, 'D'), (7, 'D', 2, 'D'), (8, 'C', 14, 'D'), (8, 'D', 8, 'D'), (9, 'C', 0, 'C'), (9, 'D', 10, 'D'), (10, 'C', 8, 'C'), (10, 'D', 15, 'C'), (11, 'C', 6, 'D'), (11, 'D', 5, 'D'), (12, 'C', 6, 'D'), (12, 'D', 9, 'D'), (13, 'C', 9, 'D'), (13, 'D', 8, 'D'), (14, 'C', 8, 'D'), (14, 'D', 13, 'D'), (15, 'C', 4, 'C'), (15, 'D', 5, 'C')], 0, C
(\textbf{Trained FSM 3})\item Feld: 1.0, 0.5, 200
\item Firm But Fair
\item Fool Me Forever
\item Fool Me Once
\item Forgetful Fool Me Once: 0.05
\item Forgetful Grudger
\item Forgiver
\item Forgiving Tit For Tat
(\textbf{FTfT})\item Fortress3
\item Fortress4
\item GTFT: 0.33
\item General Soft Grudger: n=1,d=4,c=2
\item Gradual
\item Gradual Killer: ('D', 'D', 'D', 'D', 'D', 'C', 'C')
\item Grofman
\item Grudger
\item GrudgerAlternator
\item Grumpy: Nice, 10, -10
\item Handshake
\item Hard Go By Majority
\item Hard Go By Majority: 10
\item Hard Go By Majority: 20
\item Hard Go By Majority: 40
\item Hard Go By Majority: 5
\item Hard Prober
\item Hard Tit For 2 Tats
(\textbf{HTf2T})\item Hard Tit For Tat
(\textbf{HTfT})\item Hesitant QLearner
\item Hopeless
\item Inverse
\item Inverse Punisher
\item Joss: 0.9
\item Level Punisher
\item Limited Retaliate 2: 0.08, 15
\item Limited Retaliate 3: 0.05, 20
\item Limited Retaliate: 0.1, 20
\item MEM2
\item Math Constant Hunter
\item Meta Hunter Aggressive: 7 players
\item Meta Hunter: 6 players
\item Naive Prober: 0.1
\item Negation
\item Nice Average Copier
\item Nydegger
\item Omega TFT: 3, 8
\item Once Bitten
\item Opposite Grudger
\item PSO Gambler 1\_1\_1
\item PSO Gambler 2\_2\_2
\item PSO Gambler 2\_2\_2 Noise 05
\item PSO Gambler Mem1
\item Predator
\item Prober
\item Prober 2
\item Prober 3
\item Prober 4
\item Pun1
\item Punisher
\item Raider
\item Random Hunter
\item Random: 0.5
\item Remorseful Prober: 0.1
\item Resurrection
\item Retaliate 2: 0.08
\item Retaliate 3: 0.05
\item Retaliate: 0.1
\item Revised Downing: True
\item Ripoff
\item Risky QLearner
\item SelfSteem
\item ShortMem
\item Shubik
\item Slow Tit For Two Tats
\item Slow Tit For Two Tats 2
\item Sneaky Tit For Tat
\item Soft Go By Majority
\item Soft Go By Majority: 10
\item Soft Go By Majority: 20
\item Soft Go By Majority: 40
\item Soft Go By Majority: 5
\item Soft Grudger
\item Soft Joss: 0.9
\item SolutionB1
\item SolutionB5
\item Spiteful Tit For Tat
\item Stochastic Cooperator
\item Stochastic WSLS: 0.05
\item Suspicious Tit For Tat
\item Tester
\item ThueMorse
\item ThueMorseInverse
\item Thumper
\item Tit For 2 Tats
(\textbf{Tf2T})\item Tit For Tat
(\textbf{TfT})\item Tricky Cooperator
\item Tricky Defector
\item Tullock: 11
\item Two Tits For Tat
(\textbf{2TfT})\item VeryBad
\item Willing
\item Win-Shift Lose-Stay: D
(\textbf{WShLSt})\item Win-Stay Lose-Shift: C
(\textbf{WStLSh})\item Winner12
\item Winner21
\item Worse and Worse
\item Worse and Worse 2
\item Worse and Worse 3
\item ZD-Extort-2 v2: 0.125, 0.5, 1
\item ZD-Extort-2: 0.1111111111111111, 0.5
\item ZD-Extort-4: 0.23529411764705882, 0.25, 1
\item ZD-GEN-2: 0.125, 0.5, 3
\item ZD-GTFT-2: 0.25, 0.5
(\textbf{ZD-GTFT})\item ZD-SET-2: 0.25, 0.0, 2