### 2017-05-15 tex/list_of_players.tex

\item $\phi$
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item $\pi$
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item $e$
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item ALLCorALLD
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Adaptive
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Adaptive Pavlov 2006
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Adaptive Pavlov 2011
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Adaptive Tit For Tat: 0.5
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Aggravater
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Alternator
 - \textit{Deterministic} - \textit{Memory depth}: 1\item Alternator Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Anti Tit For Tat
 - \textit{Deterministic} - \textit{Memory depth}: 1\item AntiCycler
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Appeaser
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Arrogant QLearner
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Average Copier
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Better and Better
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Bully
 - \textit{Deterministic} - \textit{Memory depth}: 1\item Calculator
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Cautious QLearner
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item CollectiveStrategy
(\textbf{CS}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Contrite Tit For Tat
(\textbf{CTfT}) - \textit{Deterministic} - \textit{Memory depth}: 3\item Cooperator
 - \textit{Deterministic} - \textit{Memory depth}: 0\item Cooperator Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Cycle Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Cycler CCCCCD
 - \textit{Deterministic} - \textit{Memory depth}: 5\item Cycler CCCD
 - \textit{Deterministic} - \textit{Memory depth}: 3\item Cycler CCCDCD
 - \textit{Deterministic} - \textit{Memory depth}: 5\item Cycler CCD
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Cycler DC
 - \textit{Deterministic} - \textit{Memory depth}: 1\item Cycler DDC
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Davis: 10
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Defector
 - \textit{Deterministic} - \textit{Memory depth}: 0\item Defector Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Desperate
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Doubler
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item EasyGo
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Eatherley
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Eventual Cycle Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Evolved ANN
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Evolved ANN 5
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Evolved ANN 5 Noise 05
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Evolved FSM 16
 - \textit{Deterministic} - \textit{Memory depth}: 16\item Evolved FSM 16 Noise 05
 - \textit{Deterministic} - \textit{Memory depth}: 16\item Evolved FSM 4
 - \textit{Deterministic} - \textit{Memory depth}: 4\item Evolved HMM 5
 - \textit{Stochastic} - \textit{Memory depth}: 5\item EvolvedLookerUp1\_1\_1
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item EvolvedLookerUp2\_2\_2
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item FSM Player: [(0, 'C', 0, 'C'), (0, 'D', 3, 'C'), (1, 'C', 5, 'D'), (1, 'D', 0, 'C'), (2, 'C', 3, 'C'), (2, 'D', 2, 'D'), (3, 'C', 4, 'D'), (3, 'D', 6, 'D'), (4, 'C', 3, 'C'), (4, 'D', 1, 'D'), (5, 'C', 6, 'C'), (5, 'D', 3, 'D'), (6, 'C', 6, 'D'), (6, 'D', 6, 'D'), (7, 'C', 7, 'D'), (7, 'D', 5, 'C')], 0, C
(\textbf{TF3}) - \textit{Deterministic} - \textit{Memory depth}: 16\item FSM Player: [(0, 'C', 13, 'D'), (0, 'D', 12, 'D'), (1, 'C', 3, 'D'), (1, 'D', 4, 'D'), (2, 'C', 14, 'D'), (2, 'D', 9, 'D'), (3, 'C', 0, 'C'), (3, 'D', 1, 'D'), (4, 'C', 1, 'D'), (4, 'D', 2, 'D'), (5, 'C', 12, 'C'), (5, 'D', 6, 'C'), (6, 'C', 1, 'C'), (6, 'D', 14, 'D'), (7, 'C', 12, 'D'), (7, 'D', 2, 'D'), (8, 'C', 7, 'D'), (8, 'D', 9, 'D'), (9, 'C', 8, 'D'), (9, 'D', 0, 'D'), (10, 'C', 2, 'C'), (10, 'D', 15, 'C'), (11, 'C', 7, 'D'), (11, 'D', 13, 'D'), (12, 'C', 3, 'C'), (12, 'D', 8, 'D'), (13, 'C', 7, 'C'), (13, 'D', 10, 'D'), (14, 'C', 10, 'D'), (14, 'D', 7, 'D'), (15, 'C', 15, 'C'), (15, 'D', 11, 'D')], 0, C
(\textbf{TF2}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item FSM Player: [(0, 'C', 7, 'C'), (0, 'D', 1, 'C'), (1, 'C', 11, 'D'), (1, 'D', 11, 'D'), (2, 'C', 8, 'D'), (2, 'D', 8, 'C'), (3, 'C', 3, 'C'), (3, 'D', 12, 'D'), (4, 'C', 6, 'C'), (4, 'D', 3, 'C'), (5, 'C', 11, 'C'), (5, 'D', 8, 'D'), (6, 'C', 13, 'D'), (6, 'D', 14, 'C'), (7, 'C', 4, 'D'), (7, 'D', 2, 'D'), (8, 'C', 14, 'D'), (8, 'D', 8, 'D'), (9, 'C', 0, 'C'), (9, 'D', 10, 'D'), (10, 'C', 8, 'C'), (10, 'D', 15, 'C'), (11, 'C', 6, 'D'), (11, 'D', 5, 'D'), (12, 'C', 6, 'D'), (12, 'D', 9, 'D'), (13, 'C', 9, 'D'), (13, 'D', 8, 'D'), (14, 'C', 8, 'D'), (14, 'D', 13, 'D'), (15, 'C', 4, 'C'), (15, 'D', 5, 'C')], 0, C
(\textbf{TF1}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Feld: 1.0, 0.5, 200
 - \textit{Stochastic} - \textit{Memory depth}: 200\item Firm But Fair
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Fool Me Forever
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Fool Me Once
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Forgetful Fool Me Once: 0.05
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Forgetful Grudger
 - \textit{Deterministic} - \textit{Memory depth}: 10\item Forgiver
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Forgiving Tit For Tat
(\textbf{FTfT}) - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Fortress3
 - \textit{Deterministic} - \textit{Memory depth}: 3\item Fortress4
 - \textit{Deterministic} - \textit{Memory depth}: 4\item GTFT: 0.33
 - \textit{Stochastic} - \textit{Memory depth}: 1\item General Soft Grudger: n=1,d=4,c=2
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Gradual
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Gradual Killer: ('D', 'D', 'D', 'D', 'D', 'C', 'C')
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Grofman
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Grudger
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item GrudgerAlternator
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Grumpy: Nice, 10, -10
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Handshake
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Hard Go By Majority
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Hard Go By Majority: 10
 - \textit{Deterministic} - \textit{Memory depth}: 10\item Hard Go By Majority: 20
 - \textit{Deterministic} - \textit{Memory depth}: 20\item Hard Go By Majority: 40
 - \textit{Deterministic} - \textit{Memory depth}: 40\item Hard Go By Majority: 5
 - \textit{Deterministic} - \textit{Memory depth}: 5\item Hard Prober
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Hard Tit For 2 Tats
(\textbf{HTf2T}) - \textit{Deterministic} - \textit{Memory depth}: 3\item Hard Tit For Tat
(\textbf{HTfT}) - \textit{Deterministic} - \textit{Memory depth}: 3\item Hesitant QLearner
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Hopeless
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Inverse
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Inverse Punisher
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Joss: 0.9
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Level Punisher
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Limited Retaliate 2: 0.08, 15
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Limited Retaliate 3: 0.05, 20
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Limited Retaliate: 0.1, 20
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item MEM2
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Math Constant Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Meta Hunter Aggressive: 7 players
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Meta Hunter: 6 players
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Naive Prober: 0.1
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Negation
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Nice Average Copier
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Nydegger
 - \textit{Deterministic} - \textit{Memory depth}: 3\item Omega TFT: 3, 8
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Once Bitten
 - \textit{Deterministic} - \textit{Memory depth}: 12\item Opposite Grudger
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item PSO Gambler 1\_1\_1
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item PSO Gambler 2\_2\_2
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item PSO Gambler 2\_2\_2 Noise 05
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item PSO Gambler Mem1
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Predator
 - \textit{Deterministic} - \textit{Memory depth}: 9\item Prober
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Prober 2
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Prober 3
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Prober 4
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Pun1
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Punisher
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Raider
 - \textit{Deterministic} - \textit{Memory depth}: 3\item Random Hunter
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Random: 0.5
 - \textit{Stochastic} - \textit{Memory depth}: 0\item Remorseful Prober: 0.1
 - \textit{Stochastic} - \textit{Memory depth}: 2\item Resurrection
 - \textit{Deterministic} - \textit{Memory depth}: 1\item Retaliate 2: 0.08
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Retaliate 3: 0.05
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Retaliate: 0.1
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Revised Downing: True
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Ripoff
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Risky QLearner
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item SelfSteem
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item ShortMem
 - \textit{Deterministic} - \textit{Memory depth}: 10\item Shubik
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Slow Tit For Two Tats
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Slow Tit For Two Tats 2
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Sneaky Tit For Tat
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Soft Go By Majority
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Soft Go By Majority: 10
 - \textit{Deterministic} - \textit{Memory depth}: 10\item Soft Go By Majority: 20
 - \textit{Deterministic} - \textit{Memory depth}: 20\item Soft Go By Majority: 40
 - \textit{Deterministic} - \textit{Memory depth}: 40\item Soft Go By Majority: 5
 - \textit{Deterministic} - \textit{Memory depth}: 5\item Soft Grudger
 - \textit{Deterministic} - \textit{Memory depth}: 6\item Soft Joss: 0.9
 - \textit{Stochastic} - \textit{Memory depth}: 1\item SolutionB1
 - \textit{Deterministic} - \textit{Memory depth}: 3\item SolutionB5
 - \textit{Deterministic} - \textit{Memory depth}: 5\item Spiteful Tit For Tat
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Stochastic Cooperator
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Stochastic WSLS: 0.05
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Suspicious Tit For Tat
 - \textit{Deterministic} - \textit{Memory depth}: 1\item Tester
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item ThueMorse
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item ThueMorseInverse
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Thumper
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Tit For 2 Tats
(\textbf{Tf2T}) - \textit{Deterministic} - \textit{Memory depth}: 2\item Tit For Tat
(\textbf{TfT}) - \textit{Deterministic} - \textit{Memory depth}: 1\item Tricky Cooperator
 - \textit{Deterministic} - \textit{Memory depth}: 10\item Tricky Defector
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Tullock: 11
 - \textit{Stochastic} - \textit{Memory depth}: 11\item Two Tits For Tat
(\textbf{2TfT}) - \textit{Deterministic} - \textit{Memory depth}: 2\item VeryBad
 - \textit{Deterministic} - \textit{Memory depth}: \(\infty\)\item Willing
 - \textit{Stochastic} - \textit{Memory depth}: 1\item Win-Shift Lose-Stay: D
(\textbf{WShLSt}) - \textit{Deterministic} - \textit{Memory depth}: 1\item Win-Stay Lose-Shift: C
(\textbf{WSLS}) - \textit{Deterministic} - \textit{Memory depth}: 1\item Winner12
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Winner21
 - \textit{Deterministic} - \textit{Memory depth}: 2\item Worse and Worse
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Worse and Worse 2
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item Worse and Worse 3
 - \textit{Stochastic} - \textit{Memory depth}: \(\infty\)\item ZD-Extort-2 v2: 0.125, 0.5, 1
 - \textit{Stochastic} - \textit{Memory depth}: 1\item ZD-Extort-2: 0.1111111111111111, 0.5
 - \textit{Stochastic} - \textit{Memory depth}: 1\item ZD-Extort-4: 0.23529411764705882, 0.25, 1
 - \textit{Stochastic} - \textit{Memory depth}: 1\item ZD-GEN-2: 0.125, 0.5, 3
 - \textit{Stochastic} - \textit{Memory depth}: 1\item ZD-GTFT-2: 0.25, 0.5
 - \textit{Stochastic} - \textit{Memory depth}: 1\item ZD-SET-2: 0.25, 0.0, 2
 - \textit{Stochastic} - \textit{Memory depth}: 1